\plaintitle

### 2016-07-19 drafts/labor/introduction.tex
notes and revisions for the theory paper

% %!TEX root = ./trackingWork.tex
\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}
\section{Introduction}
% each sentence feels like a thought that's unrelated to the next one
The past decade has seen a proliferation in on--demand work
where workers' statuses are so fleeting that
they have been described as ``transient''.
After identifying that myriad tasks
--- from the mundane to the complex ---
can be accomplished
% beginning with ``information work'' such as on
% Amazon Mechanical Turk (AMT).
% ??? ->
% Stemming from the recognition that
% complex tasks might be achievable
% --- and indeed trivial ---
by directing and managing the ``crowd'',
the research and and industry communities have flocked to sites of labor
like Amazon's Mechanical Turk (AMT)
to solicit workers.
Researchers in particular have taken to the space in earnest,
finding opportunities to enable new forms of work
as well as using these workers (``Turkers'')
as representative populations of the public
\cite{bernsteinSoylent,redballoon,paolacci2010running}.

This approach to managing work (and in particular workers)
can be differentiated from other formulations by a constellation of traits:
\begin{inlinelist}
\item distributing the workforce,
\item distancing workers from the context of the work, and
\item compensating work on a per--task basis rather than for time
\end{inlinelist}.
These features can be found elsewhere
--- for instance,
distributed workforces
are a well--known feature of a global economy,
%
and abstracting workers from the products of their labor
is well--associated with
\citeauthor{marx2012economic} ---
but these individual components have been explored to some extent
(for instance, \citeauthor{fevre1986contract} and later \citeauthor{kalleberg2000nonstandard}),
and we focus on the newfound interest in
this particular paradigm of work
% While contract work is familiar enough,
% much of this work focuses on comprehensive work
% (for example, creative or professional work like landscaping, design, etc\dots)
\cite{kalleberg2000nonstandard,fevre1986contract,marx2012economic}.

@@

@@

Researchers have made efforts to understand the \textit{people} that have gravitated toward on--demand work
since its emergence and popularization,
but as the form of work has grown and changed, so too have the demographics of workers
\cite{Ross,whoareNOTtheTurkers}.
Some of this research has been motivated by the realization of the sociality of gig work,
and the frustration and disenfranchisement that these systems embody
\cite{turkopticon,dynamo}.
Other work has focused on the outcomes of work,
reflecting on the resistance workers express against digitally mediated labor markets
\cite{uberAlgorithm}.
Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
workers increasingly expressing frustration with the disenfranchisement of the work--flows pioneered by researchers in the decade leading up till now,
and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.

>>>>

% %!TEX root = ./trackingWork.tex
\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\newcommand{\note}[1]{\onlyinsubfile{\textit{#1}}}

\section{Introduction}
% each sentence feels like a thought that's unrelated to the next one
The past decade has seen a proliferation in on--demand work
where workers' statuses are so fleeting that
they have been described as ``transient''.
After identifying that myriad tasks
--- from the mundane to the complex ---
can be accomplished
% beginning with ``information work'' such as on
% Amazon Mechanical Turk (AMT).
% ??? ->
% Stemming from the recognition that
% complex tasks might be achievable
% --- and indeed trivial ---
by directing and managing the ``crowd'',
the research and and industry communities have flocked to sites of labor
like Amazon's Mechanical Turk (AMT)
to solicit workers.
Researchers in particular have taken to the space in earnest,
finding opportunities to enable new forms of work
as well as using these workers (``Turkers'')
as representative populations of the public
\cite{bernsteinSoylent,redballoon,paolacci2010running}.

\note{Is this different from other contract work? yes. here's why:}

This approach to managing work (and in particular workers)
can be differentiated from other formulations by a constellation of traits:
\begin{inlinelist}
\item distributing the workforce, % gray
\item distancing workers from the context of the work, and
\item compensating work on a per--task basis rather than for time
% citations! future of crowdwork or mary's or lily's work / quin and bederson
\end{inlinelist}.
These features can be found elsewhere
--- for instance,
distributed workforces
are a feature of global labor markets,
and abstracting workers from the products of their labor
is well--associated with
\citeauthor{marx2012economic} ---
but these components have been explored only individually
(for instance, by \citeauthor{fevre1986contract} and later \citeauthor{kalleberg2000nonstandard}),
and we focus on the newfound interest in
work arranging these components in what holistically seems to be a new mode of work
% While contract work is familiar enough,
% much of this work focuses on comprehensive work
% (for example, creative or professional work like landscaping, design, etc\dots)
\cite{kalleberg2000nonstandard,fevre1986contract,marx2012economic}.

\note{why is this important? because it's bigger than AMT}

@@

\note{this is getting into more general ``how has this moved forward'' kind of work. save for later?}

@@

Researchers have made efforts to understand the \textit{people} that have gravitated toward on--demand work
since its emergence and popularization,
but as the form of work has grown and changed, so too have the demographics of workers
\cite{Ross,whoareNOTtheTurkers}.
Some of this research has been motivated by the realization of the sociality of gig work,
and the frustration and disenfranchisement that these systems embody
\cite{turkopticon,dynamo}.
Other work has focused on the outcomes of work,
reflecting on the resistance workers express against digitally mediated labor markets
\cite{uberAlgorithm}.

Bringing this broader body of work together, perhaps we can think of this as documenting the ``fallout'' of on--demand labor;
workers increasingly expressing frustration with the disenfranchisement of the work--flows pioneered by researchers in the decade leading up till now,
and researchers beginning to appreciate the unforeseen costs of transient labor and the impacts both on the work and the workers.

\note{this probably doesn't belong here but i need to put it somewhere!}

From various perspectives and with an eye toward myriad stakeholders,
this body of work has ostensibly sought to answer one overarching question:
What does the future hold for work and those that do it? % expand on ...?
% make it clearer
% inevitable/predictable outcomes?
% aspirational -> how can we make work better for people
Whether that research has attempted to articulate the potential to achieve difficult and creative tasks,
or wrestled with the frustration that workers face on a near--constant basis,
these roads of inquiry have, it seems, looked to the horizon to try to predict where we are heading.

Rather than looking necessarily to the creation of systems
(both as genuine agents and as provocateurs (e.g. Soylent and Turkopticon)),
we turn to the \textit{past}, uncovering ``piecework'' as
a useful model for understanding the mode of work that has captured the imagination of entrepreneurs and researchers.
